---
title: "Emissions and Perceptions of Climate Change"
author: "Rae Dunbar"
date: "11/13/2022"
output: html_document
---

## Introduction

In the great pursuit of mitigating climate change, countries have differed greatly in their strategies and successes. Denmark has one of the most ambitious mitigation plans in the world pledging to cut 70% of its emissions by 2030.^[Etsy, D. C., "Why Denmark Wants to Be a ‘Frontrunner’ in the Fight Against Climate Change." Yale School of the Environment.2021.] As other countries trail behind, illuminating how Denmark is successfully reducing its emissions is pertinent for other countries. Although reducing emissions involves extensive effort from multiple sectors, the citizens also greatly determine the success of mitigation plans.^[Weber, E. U., "What Shapes perceptions of climate change?," WIRES Climate Change.2010] What determines people's perceptions and willingness to decrease their own emissions, is therefore important for countries to understand. To delve into this question data from two global surveys on climate perceptions will be analyzed along with green house gas emissions and the reduction rate of countries. Special emphasis will be on comparing Denmark due to the countrie's success, and the United States, due to it's world’s leading production of oil and natural gas.^[Gross, S., "The United States can take climate change seriously while leading the world in oil and gas production." Policy 2022 Brookings. 2022.] 

This data analysis will consist of three sections: the greenhouse gas emissions of the U.S. and Denmark, the climate change perceptions of their citizens, and the correlation between emissions and climate change perceptions, using data from many countries.

In the emissions section Denmark's the United State's per capita emissions and the rate of decrease will both be compared.We hypothesize that Denmark will have less emissions and a faster emission reduction rate than the United States, because the U. S. had the highest cumulative greenhouse gas emissions from 850-2021 whereas Denmark has one of the most ambitious emission reduction plans in the world.^[Evans, S., "Analysis: Which countries are historically responsible for climate change?." Carbon Brief, 2021.]

The second section will include an analysis of several quesitons regarding climate change for both the U.S. and Denmark. These will whether the citizens are aware of climate change, how worried they are, and how big of a threat they believe climate change is in the next 20 years. We hypothesize Denmark will have a significantly higher percentage of people who are aware of climate change than the U.S, because Denmark hosted COP15, which sparked national social media attention on climate change, exposing many Danes to the concept.^[Gunster, Shane. “Covering Copenhagen: Climate Change in BS Media.” Canadian Journal of Communication 36, no.3 (Nov 2011): 477-502. DOI: 10.22230/cjc.2011v36n3a2367] We further hypothesize that more of Denmark's citizens will be very worried about climate change and see it as an extreme threat than U. S. citizens, because because climate change is uniformly depicted in Danish media as a problem, but it is not in the U.S.^4

In the third section we will overlay the survey data with the emissions data to deterrmine the percentage of causation. We predict that people’s perceptions of climate change impacts about 30% of greenhouse gas emissions of a country, given there are several other significantly related variables, such as population, income, and political structure.^[Wang, Q., Feng, G., Wang, H., Chang. C., " The influence of political ideology on greenhouse gas emissions" Global Environmental Change 74, no.1 (May 2022):102496. DOOI: https://doi.org/10.1016/j.gloenvcha.2022.102496.]  

## Statistcal Analysis and Results

Statistical analysis will consist of the aforementioned three sections.

## Climate Change Emissions of Countries

Preparing R
```{r}
rm(list=ls())
library(tidyverse)
library(here)
library(ggfortify)
```

Pulling in the corrected Emissions data. For the raw data see the raw data folder on github. Please also see the metadata txt files per each data set in the metadata folder.

Lets first view the emissions of Denmark and the United States.

Reading in the Data
```{r}
emissions<-read.csv(here("Data for the Code", "Corrected_Data_sets", "emissions_corrected_US_Denmark.csv"), stringsAsFactors = TRUE)
```

### Creating the figure

???HOW DO I GET RID OF THE LINE?
```{r}

#year <- function(x) as.POSIXlt(x)$year + 1900
# I do not know how the above code does this :(
ggplot(emissions, aes(year, emissions)) + 
  geom_path() + geom_point(aes(colour=country))+ theme_bw() 
 
```
### Running the Test
```{r}

```

### Rate of Emissions

To determine rate the emissions must be divided by the year. 

???? HOW DO THAT WITH TIME SERIES???


## Perceptions of Climate Change



## Summary Discussion

## Challenges and Limitations
